### 2017-11-20 README.Rmd
use_analysis()

---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
```

# refractoryANA

The goal of refractoryANA is to ...

### Citation

Please cite this compendium as:

> W. Francuzik, M. Worm, (`r format(Sys.Date(), "%Y")`). _Risk Factors and Managment of Refractory Anaphylaxis Research Compendium_. Accessed `r format(Sys.Date(), "%d %b %Y")`. Online at <https://doi.org/xxx/xxx>

### Installation

You can install refractoryANA from github with:

```{r gh-installation, eval = FALSE}
# install.packages("devtools")
devtools::install_github("wolass/refractoryANA")
```

### Licenses

**Text and figures :**  [CC-BY-4.0](http://creativecommons.org/licenses/by/4.0/)

**Code :** See the [DESCRIPTION](DESCRIPTION) file

**Data :** [CC-0](http://creativecommons.org/publicdomain/zero/1.0/) attribution requested in reuse

### Contributions

We welcome contributions from everyone. Before you get started, please see our [contributor guidelines](CONTRIBUTING.md). Please note that this project is released with a [Contributor Code of Conduct](CONDUCT.md). By participating in this project you agree to abide by its terms.